---
title: "Final Code Preparing Data"
author: "Michael Latimer, Angela Adamaris Gutierrez Castillo, Ojashvee Gupta, Marlon Brooks, Shanice Sinclair"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data <- read.csv("Spotify_Master_US_Data.csv", stringsAsFactors = F)
library(tm)
library(SnowballC)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggthemes)
```

## Converting Lyrics to Corpus and Inspecting Song

```{r}
corpus <- Corpus(VectorSource(data$Lyrics))

corpus[[52]][1]
```

## Convert text to lower case

```{r, echo=FALSE}
corpus <- tm_map(corpus, FUN = content_transformer(tolower))

corpus[[240]][1]
```

## Remove Punctuation

```{r, echo=FALSE}

corpus <- tm_map(corpus,FUN = removePunctuation)

corpus[[52]][1]
```

## Remove \n where computer is denoting new line

```{r}

corpus = tm_map(corpus,
                FUN = content_transformer(FUN = function(x)gsub(pattern = '\n',
                                                                replacement = ' ',x = x)))

corpus[[52]][1]
```

## Remove Stop Words

```{r}

corpus <- tm_map(corpus, FUN = removeWords,c(stopwords('english')))

corpus[[52]][1]

```

# Remove Stop Words in Spanish

```{r}

corpus <- tm_map(corpus, FUN = removeWords,c(stopwords('spanish')))

corpus[[3]][1]

```

# Strip White Space

```{r}

corpus <- tm_map(corpus,FUN = stripWhitespace)

corpus[[52]][1]

```

# Stem Words

```{r}
corpus <- tm_map(corpus,FUN = stemDocument)

corpus[[52]][1]
```

# Add Cleaned Corpus to Original Data Set

```{r}

data$CleanText <- get("content", corpus)

str(data)

```

# Create Cleaned CSV

```{r}

creationFile = data.frame(ID = data$ID, Position = data$Position, Title = data$Title, Artist = data$Artist, Year = data$Year, Streams = data$Streams, Lyrics = data$Lyrics, CleanText = data$CleanText)

write.csv(creationFile,'Spotify_Master_US_Data_Cleaned.csv', row.names=FALSE)

```

# Preliminary Exploring - Create Dictionary

```{r}

dict = findFreqTerms(DocumentTermMatrix(Corpus(VectorSource(data$Lyrics))),
                     lowfreq = 0)
dict_corpus = Corpus(VectorSource(dict))
```

# Create Document Term Matrix and Inspect

```{r}

dtm <- DocumentTermMatrix(corpus)

inspect(dtm[52,"ever"])

```

# Remove Sparse Terms and Complete Stems

```{r}

xdtm = removeSparseTerms(dtm,sparse = 0.95)

xdtm = as.data.frame(as.matrix(xdtm))
colnames(xdtm) = stemCompletion(x = colnames(xdtm),
                                dictionary = dict_corpus,
                                type='prevalent')
colnames(xdtm) = make.names(colnames(xdtm))

```

# Review Frequency of Tokens

```{r}

sort(colSums(xdtm),decreasing = T)

```

